SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python

### 2019-05-25 scipy-1.0/paper.tex
Minor punctuation/reference tweaks

SciPy's arrival at this point is surprising, and somewhat anomalous.
When started, the library had little funding, and was written mainly
by graduate students---many of them without a computer science education, and often without the
blessing of their advisors.  To even imagine that a small group of
``rogue'' student programmers could upend the already well-established
ecosystem of research software---backed by millions in funding, and
many hundreds of highly qualified engineers\cite{mathworks-globe-97,
esri-revenue,bloom-wolfram}---was preposterous.

@@

In 1995, Jim Hugunin, a graduate student from MIT, wrote the first
message in a new Python Matrix Special Interest Group (Matrix-SIG)
mailing list\cite{Hugunin-first}:
\begin{quote}
There seems to be a fair amount of interest in the Python community
concerning the addition of numeric operations to Python.  My own desire is
to have as large a library of matrix based functions available as possible
(linear algebra, eigenfunctions, signal processing, statistics, etc.).  In
order to ensure that all of these libraries interoperate, there needs to
be agreement on a basic matrix object that can be used to represent arrays
of numbers.
\end{quote}
Over the next several months, conversations on that mailing
list by, among others, Jim Fulton, Jim Hugunin, Paul Dubois, Konrad
Hinsen, and Guido van Rossum, led to the creation of a package called Numeric with an array object
that supported a high number of dimensions.  Jim Hugunin explained the utility
of Python for numerical computation\cite{Hugunin-whitepaper}:
\begin{quote}
I've used almost all of the available numerical languages at one time
or another over the past 8 years. One thing I've noticed is that over
time, the designers of these languages are steadily adding more of the
features that one would expect to find in a general-purpose
programming language.
\end{quote}
This is still a distinguishing feature of Python for science, and one of the
reasons why it has been so successful in the realm of data science: instead of
adding general features to a language designed for numerical and scientific
computing, here scientific features are added to a general purpose language.  
This broadens the scope of problems that can be addressed easily, expands the
sources of data that are readily accessible, and increases the size of the
community that develops code for the platform.

@@

By the middle to late 2000s, SciPy was starting to mature after a long phase of significant
growth and adoption.
The informal workshops grew into international conferences with many
hundreds of attendees. Special issues were organized and published in a
leading scientific journal\cite{dubois2007guest}, and the scope of the SciPy library
narrowed, while the breadth of the ecosystem grew
through a new type of auxiliary package, the scikit\cite{scikits-general}. 
The tooling, development, and release processes became more professional.
SciPy was expanded carefully, with the patience affordable in open source 
projects and via best practices common in industry \cite{millman2014developing}.

@@

\subsection*{SciPy Today}
At the time of writing, the SciPy library consists of nearly 
600,000 lines of code organized in 16 subpackages. 
Over 85,000 GitHub repositories and almost 5000 packages depend
on SciPy. Some of the major
feature highlights from the three years preceding
SciPy 1.0 are discussed in Section \ref{sec:technical_improvements}, 
and milestones in its history are highlighted in Figure~\ref{fig:timeline}.
% At the time of writing, SciPy produces a new release
% approximately every six months, and some of the major
% feature highlights from the three years preceding
% SciPy 1.0 are discussed in a technical improvements
% section below. Some longer-term milestone events are
% highlighted in Figure~\ref{fig:timeline}.

\begin{figure}[H]
\centering
\includegraphics[width=0.95\textwidth]{static/scipy_timeline}
\caption{Major milestones from SciPy's initial release in 2001 to
the release of SciPy 1.0 in 2017. Logos reprinted with permission.}
\label{fig:timeline}
\end{figure}

@@

Python, Cython, Fortran, C and C++ are the programming languages used to
implement scientific algorithms in the SciPy library. An analysis of our code
base using the \texttt{linguist} library\cite{linguistref} provides a 
detailed breakdown as \% composition by programming language in 
SciPy (Table~\ref{tab:linguist}).

@@

Cython has been described as a creole language that mixes the best parts of Python and
lower-level C / C++ paradigms\cite{behnel2011cython}. We often use Cython
as a glue between well-established low-level scientific computing libraries written
in C / C++ and the Python interface offered by SciPy. We also use Cython to enable performance
enhancements in Python code, especially for cases where heavily used inner
loops benefit from a compiled code with static typing.

@@

From a new features standpoint, \texttt{scipy.sparse} matrices and linear
operators now support the Python matrix multiplication (@) operator when
available. We've added \texttt{scipy.sparse.norm} and
\texttt{scipy.sparse.random} for computing sparse matrix norms and drawing
random variates from arbitrary distributions, respectively. Also, we've made a
concerted effort to bring the \texttt{scipy.sparse} API into line with the
equivalent NumPy API where possible.

@@

The SciPy test suite is orchestrated by a continuous integration matrix that
includes POSIX and Windows (32/64-bit) platforms managed by Travis CI and
AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
unit tests in the test suite, which is written for usage with the \texttt{pytest}
framework, and with 87 \% and 45 \% line coverages for Python and compiled
code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}). Documentation for the code is automatically built and hosted by
the CircleCI service to facilitate evaluation of documentation changes /
integrity.  Our full test suite also passes with PyPy3, a just-in-time compiled
version of the Python language.

>>>>

SciPy's arrival at this point is surprising and somewhat anomalous.
When started, the library had little funding, and was written mainly
by graduate students---many of them without a computer science education, and often without the
blessing of their advisors.  To even imagine that a small group of
``rogue'' student programmers could upend the already well-established
ecosystem of research software---backed by millions in funding and
many hundreds of highly qualified engineers\cite{mathworks-globe-97,
esri-revenue,bloom-wolfram}---was preposterous.

@@

In 1995, Jim Hugunin, a graduate student from MIT, wrote the first
message in a new Python Matrix Special Interest Group (Matrix-SIG)
mailing list\cite{Hugunin-first}:
\begin{quote}
There seems to be a fair amount of interest in the Python community
concerning the addition of numeric operations to Python.  My own desire is
to have as large a library of matrix based functions available as possible
(linear algebra, eigenfunctions, signal processing, statistics, etc.).  In
order to ensure that all of these libraries interoperate, there needs to
be agreement on a basic matrix object that can be used to represent arrays
of numbers.
\end{quote}
Over the next several months, conversations on that mailing
list by, among others, Jim Fulton, Jim Hugunin, Paul Dubois, Konrad
Hinsen, and Guido van Rossum led to the creation of a package called Numeric with an array object
that supported a high number of dimensions.  Jim Hugunin explained the utility
of Python for numerical computation\cite{Hugunin-whitepaper}:
\begin{quote}
I've used almost all of the available numerical languages at one time
or another over the past 8 years. One thing I've noticed is that over
time, the designers of these languages are steadily adding more of the
features that one would expect to find in a general-purpose
programming language.
\end{quote}
This is still a distinguishing feature of Python for science, and one of the
reasons why it has been so successful in the realm of data science: instead of
adding general features to a language designed for numerical and scientific
computing, here scientific features are added to a general purpose language.  
This broadens the scope of problems that can be addressed easily, expands the
sources of data that are readily accessible, and increases the size of the
community that develops code for the platform.

@@

By the middle to late 2000s, SciPy was starting to mature after a long phase of significant
growth and adoption.
The informal workshops grew into international conferences with many
hundreds of attendees. Special issues were organized and published in a
leading scientific journal\cite{dubois2007guest}, and the scope of the SciPy library
narrowed, while the breadth of the ecosystem grew
through a new type of auxiliary package: the scikit\cite{scikits-general}.
The tooling, development, and release processes became more professional.
SciPy was expanded carefully, with the patience affordable in open source 
projects and via best practices common in industry \cite{millman2014developing}.

@@

\subsection*{SciPy Today}
At the time of writing, the SciPy library consists of nearly 
600,000 lines of code organized in 16 subpackages. 
Over 85,000 GitHub repositories and almost 5000 packages depend
on SciPy. Some of the major
feature highlights from the three years preceding
SciPy 1.0 are discussed in the Key Technical Improvements section below,
and milestones in its history are highlighted in Figure~\ref{fig:timeline}.
% At the time of writing, SciPy produces a new release
% approximately every six months, and some of the major
% feature highlights from the three years preceding
% SciPy 1.0 are discussed in a technical improvements
% section below. Some longer-term milestone events are
% highlighted in Figure~\ref{fig:timeline}.

\begin{figure}[H]
\centering
\includegraphics[width=0.95\textwidth]{static/scipy_timeline}
\caption{Major milestones from SciPy's initial release in 2001 to
the release of SciPy 1.0 in 2017.}
\label{fig:timeline}
\end{figure}

@@

Python, Cython, Fortran, C and C++ are the programming languages used to
implement scientific algorithms in the SciPy library. An analysis of our codebase
using the \texttt{linguist} library\cite{linguistref} provides a
detailed breakdown of SciPy's composition (Table~\ref{tab:linguist}).

@@

Cython has been described as a creole language that mixes the best parts of Python and
lower-level C / C++ paradigms\cite{behnel2011cython}. We often use Cython
as a glue between well-established, low-level scientific computing libraries written
in C / C++ and the Python interface offered by SciPy. We also use Cython to enable performance
enhancements in Python code, especially for cases where heavily used inner
loops benefit from a compiled code with static typing.

@@

From a new features standpoint, \texttt{scipy.sparse} matrices and linear
operators now support the Python matrix multiplication (@) operator.
We've added \texttt{scipy.sparse.norm} and
\texttt{scipy.sparse.random} for computing sparse matrix norms and drawing
random variates from arbitrary distributions, respectively. Also, we've made a
concerted effort to bring the \texttt{scipy.sparse} API into line with the
equivalent NumPy API where possible.

@@

The SciPy test suite is orchestrated by a continuous integration matrix that
includes POSIX and Windows (32/64-bit) platforms managed by Travis CI and
AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
unit tests in the test suite, which is written for usage with the \texttt{pytest}
framework, and with 87 \% and 45 \% line coverages for Python and compiled
code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}). Documentation for the code is automatically built and hosted by
the CircleCI service to facilitate evaluation of documentation changes /
integrity.  Our full test suite also passes with PyPy3\cite{Bolz:2009:TMP:1565824.1565827}, a just-in-time compiled
version of the Python language.

### 2019-05-25 scipy-1.0/subpackages.tex
Minor punctuation/reference tweaks

    interpolators, and wrappers for FITPACK and DFITPACK functions.

@@

    A collection of functions and classes for reading and writing MATLAB, IDL,
    Matrix Market, Fortran, NetCDF, Harwell-Boeing, WAV and ARFF data files. 

@@

    library ODRPACK.

@@

    convolution and correlation; splines; filtering and filter design;
    continuous and discrete time linear systems; waveform generation;
    window functions; wavelet computations; peak finding; and spectral

>>>>

    interpolators, and wrappers for FITPACK\cite{Dierckx:1993:CSF:151103} and DFITPACK functions.

@@

    A collection of functions and classes for reading and writing MATLAB\cite{matlab}, IDL,
    Matrix Market\cite{boisvert1997matrix}, Fortran, NetCDF\cite{rew1990netcdf},
		Harwell-Boeing\cite{harwellboeing}, WAV and ARFF data files.

@@

    library ODRPACK\cite{ODRPACK_Boggs}.

@@

    convolution and correlation, splines, filtering and filter design,
    continuous and discrete time linear systems, waveform generation,
    window functions, wavelet computations, peak finding, and spectral
